### 2014-02-20 response/candc.tex
gathering info for lit review section

\documentclass[10pt,letterpaper,draft]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\author{Michael D. Brothers}
\title{Comparing and contrasting with prior work in the field}
\begin{document}
\section{Abstract of the reviewed work}

In order to maintain the quadratic convergence properties of Newton's method in
quasi-static non-linear analysis of solid structures it is crucial to obtain
accurate, algorithmically consistent tangent-stiffness matrices. A goal of the
study described in this paper was to establish the suitability of an
under-explored method for numerical computation of tangent-stiffness operators,
referred to as ``complex-step'', and compare the method with other techniques
for numerical derivative calculation: automatic differentiation, forward
finite-difference, and central finite-difference. The complex-step method was
implemented in a massively parallel computational peridynamics code for the
purpose of this comparison.  The methods were compared through invasive
profiling of the code for accuracy, speed, efficiency, and parallel
scalability. The research provides data that can serve as practical guide for
code developers and analysts faced with choosing which method best suits the
needs of their application code.  Additionally, motivated by the reproducible
research movement, all the of the code, examples, and workflow to regenerate
the data and figures in this paper are provided as open source.

\section{Re: W. Squire and G. Trapp (1998)}
\subsection{Abstract of prior work in field}

A method to approximate derivatives of real functions using complex variables
which avoids the subtractive cancellation errors inherent in the classical
derivative approximations is described. Numerical examples illustrating the
power of the approximation are presented.

\subsection{Aspects of CS investigated}

Establishing the accuracy of the method for computing first derivatives of
real functions through empirical tests, versus forward and central difference 
methods.

\subsection{Similarities with review work}

The reviewed work uses the numerical differentiation method as shown in the
prior work. The accuracy of CS demonstrated in the prior work served as encouragement
for investigating CS further.

\subsection{Differences with review work}

In the reviewed work the CS method is used to compute derivatives of a
non-linear vector function of vector variables wrt the components of the vector
variables, an extension of the method shown in the prior work.
Additionally, automatic differentiation is compared to CS. Beyond that, not
only is the accuracy of the method measured, but its impact as part of a Newton
Raphson solution scheme for a Peridynamics simulation run on a parallel
computer is investigated.

\subsection{Novelty of reviewed work}
The reviewed work examines applications of the CS method beyond taking derivatives alone,
whereas the prior work does not.

\section{Re: A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta \emph{n.d. for local...} (2000)}
\subsection{Abstract of prior work in field}

In this paper, numerical differentiation is applied to integrate plastic
constitutive laws and to compute the corresponding consistent tangent
operators. The derivatives of the constitutive equations are approximated by
means of difference schemes. These derivatives are needed to achieve quadratic
convergence in the integration at Gauss-point level and in the solution of the
boundary value problem.  Numerical differentiation is shown to be a simple,
robust and competitive alternative to analytical derivatives. Quadratic
convergence is maintained, provided that adequate schemes and stepsizes are
chosen. This point is illustrated by means of some numerical examples.

\subsection{Aspects of CS investigated}

The prior work applies CS in computing derivatives needed for integrating
plastic stress, obtaining algorithmically consistent elastoplastic modulii, and
obtaining the Jacobian of the residual, for FEA simulations of two plastic
material models.

\subsection{Similarities with review work}

First order CS is being used to compute the Jacobian of the residual in a solid mechanics.

\subsection{Differences with review work}

The prior work demonstrates CS on a serial computer, and with a classical solid mechanics
formulation. Automatic differentiation is not investigated, and speed measurements are
not taken. Scalability studies with a variety of equivalent, progressively more refined meshes
are not done. Scalability as a function of distributed system size is not investigated.

\subsection{Novelty of reviewed work}

While the prior work does apply CS to computing Jacobians of the residual, and further applies CS
to computing elastoplastic modulii, the reviewed work is novel by contrast in offering discussion
of CS as applied to Peridynamics on a parallel computer, a larger range of problem parameters, 
CPU time data, and comparison to automatic differentiation as a competing method.

\section{Re: A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta \emph{n.d. for nontrivial...}(2000)}
\subsection{Abstract of prior work in field}

In Reference [1] the authors have shown that numerical differentiation is a
competitive alternative to analytical derivatives for the computation of
consistent tangent matrices. Relatively simple models were treated in that
reference. The approach is extended here to a complex model: the MRS-Lade model
[2,3]. This plastic model has a cone-cap yield surface and exhibits strong
coupling between the flow vector and the hardening moduli. Because of this,
derivating these quantities with respect to stresses and internal variables
—the crucial step in obtaining consistent tangent matrices— is rather involved.
Numerical differentiation is used here to approximate these derivatives. The
approximated derivatives are then used 1) to compute consistent tangent
matrices (global problem) and 2) to integrate the constitutive equation at each
Gauss point (local problem) with the Newton-Raphson method. The choice of the
stepsize (i.e. the perturbation in the approximation schemes), based on the
concept of relative stepsize, poses no difficulties. In contrast to previous
approaches for the MRS-Lade model, quadratic convergence is achieved, for both
the local and the global problems. The computational efficiency (CPU time) and
robustness of the proposed approach is illustrated by means of several
numerical examples, where the major relevant topics are discussed in detail.

\subsection{Aspects of CS investigated}

The prior work applies CS in computing derivatives needed for integrating
plastic stress, obtaining algorithmically consistent elastoplastic modulii, and
obtaining the Jacobian of the residual, for FEA simulations of the MRS-Lade material
model.

\subsection{Similarities with review work}

CS is used to compute the Jacobian of the residual. CPU runtime is measured.

\subsection{Differences with review work}

The prior work demonstrates CS on a serial computer, and with a classical solid
mechanics formulation. Automatic differentiation is not investigated.
Scalability studies with a variety of equivalent, progressively more refined
meshes are not done. Scalability as a function of distributed system size is
not investigated.

\subsection{Novelty of reviewed work}

While the prior work does apply CS to computing Jacobians of the residual, and further applies CS
to computing elastoplastic modulii, the reviewed work is novel by contrast in offering discussion
of CS as applied to Peridynamics on a parallel computer, a larger range of problem parameters, 
and comparison to automatic differentiation as a competing method.

\section{Re: J. Martins, P. Sturdza, J. Alonso (2003)}
\subsection{Abstract of prior work in field}

The complex-step derivative approximation and its application to numerical
algorithms are presented. Improvements to the basic method are suggested that
further increase its accuracy and robustness and unveil the connection to
algorithmic differentiation theory. A general procedure for the implementation
of the complex-step method is described in detail and a script is developed
that automates its implementation. Automatic implementations of the
complex-step method for Fortran and C/C++ are presented and compared to
existing algorithmic differentiation tools. The complex-step method is tested
in two large multidisciplinary solvers and the resulting sensitivities are
compared to results given by finite differences. The resulting sensitivities
are shown to be as accurate as the analyses. Accuracy, robustness, ease of
implementation and maintainability make these complex-step derivative
approximation tools very attractive options for sensitivity analysis.

\subsection{Aspects of CS investigated}

The prior work demonstrates that the CS method can be performed by using complex
datatypes for the variables of functions comprising an analysis code. The analysis code
can comprise a simulation or similar, and derivatives of response variables called
sensitivities can be computed. The accuracy of this approach is compared to automatic 
differentiation.

\subsection{Similarities with review work}

CS is used to compute derivatives.

\subsection{Differences with review work}

The prior work mentions employing CS within an iterative solver code for a sensitivity analysis, but
does not mention if CS is also used by the iterative solver itself.

\subsection{Novelty of reviewed work}

\section{Re: W. K. Anderson, J. C. Newman, D. l. Whitfield and E. J. Nielsen (2001)}
\subsection{Abstract of prior work in field}

The use of complex variables for determining sensitivity derivatives for
turbulent flows is examined. Although a step size parameter is required, the
numerical derivatives are not subject to subtractive cancellation errors and,
therefore, exhibit true second-order accuracy as the step size is reduced. As a
result, this technique guarantees two additional digits of accuracy each time
the step size is reduced one order of magnitude. This behavior is in contrast
to the use of finite differences, which suffer from inaccuracies due to
subtractive cancellation errors. In addition, the complex-variable procedure is
easily implemented into existing codes.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{A. H. Al-Mohy and N. J. Higham (2010)}
\subsection{Abstract of prior work in field}

We show that the Fréchet derivative of a matrix function f at A in the
direction E, where A and E are real matrices, can be approximated by Im f (A +
ihE)/ h for some suitably small h. This approximation, requiring a single
function evaluation at a complex argument, generalizes the complex step
approximation known in the scalar case. The approximation is proved to be of
second order in h for analytic functions f and also for the matrix sign
function. It is shown that it does not suffer the inherent cancellation that
limits the accuracy of finite difference approximations in floating point
arithmetic. However, cancellation does nevertheless vitiate the approximation
when the underlying method for evaluating f employs complex arithmetic.  The
ease of implementation of the approximation, and its superiority over finite
differences, make it attractive when specialized methods for evaluating the
Fréchet derivative are not available, and in particular for condition number
estimation when used in conjunction with a block 1-norm estimation algorithm.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{W. Jin, B. H. Dennis, and B. P. Wang (2010)}
\subsection{Abstract of prior work in field}

The semi-analytical method (SAM) is a computationally efficient and easy to
implement approach often used for the sensitivity analysis of finite element
models.  However, it is known to exhibit serious inaccuracy for shape
sensitivity analysis for structures modeled by beam, frame, plate, or shell
elements. In the present paper, we use a semi-analytical approach based on
complex variables (SACVM) to compute the sensitivity of finite element mod- els
composed of beam and plate elements. The SACVM combines the complex variable
method (CVM) with the semi-analytical method (SAM) to obtain the response
sensitivity accurately and efficiently. The current approach maintains the
computational efficiency of the semi-analytical method but with higher
accuracy. In addition, the current approach is insensitive to the choice of
step size, a feature that simplifies its use in practical problems. The method
is applicable to any structural elements including beam, frame, plate, or shell
elements and only requires minor modifications to existing finite element
codes.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{A. Voorhees, H. Millwater, and R. Bagley (2011)}
\subsection{Abstract of prior work in field}

Shape sensitivity analysis of finite element models is useful for structural
optimization and design modifications. Complex variable methods for shape
sensitivity analysis have some potential advantages over other methods. In
particular, for first order sensitivities using the complex Taylor series
expansion method (CTSE), the implementation is straightforward, only requiring
a perturbation of the finite element mesh along the imaginary axis. That is,
the real valued coordinates of the mesh are unaltered and no other
modifications to the software are required. Fourier differentiation (FD)
provides higher order sensitivities by conducting an FFT analysis of multiple
complex variable analyses around a sampling radius in the complex plane.
Implementation of complex variable sensitivity methods requires complex
variable finite element software such that complex nodal coordinates can be
used to implement a perturbation in the shape of interest in the complex
domain. All resulting finite element outputs such as displacements, strains and
stresses become complex and accurate derivatives of all finite element outputs
with respect to the shape parameter of interest are available. The
methodologies are demonstrated using two-dimensional finite element models of
linear elasticity problems with known analytical solutions. It is found that
the error in the sensitivities is primarily defined by the error in the finite
element solution not the error in the sensitivity method. Hence, more accurate
sensitivities can be obtained through mesh refinement.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{K.-L. Lai and J. L. Crassidis (2008)}
\subsection{Abstract of prior work in field}

A general framework for the first and second complex-step derivative
approximation to compute numerical derivatives is presented.  For first
derivatives the complex-step approach does not suffer roundoff errors as in
standard numerical finite-difference approaches.  Therefore, since an
arbitrarily small step size can be chosen, the complex-step approach can
achieve near analytical accuracy.  However, for second derivatives straight
implementation of the complex-step approach does suffer from roundoff errors.
Therefore, an arbitrarily small step size cannot be chosen. In this paper the
standard complex-step approach is expanded by using general complex-step sizes
to provide a wider range of accuracy for both the first- and second-derivative
approximations. Even higher accuracy formulations are obtained by repetitively
applying Richardson extrapolations. The new extensions can allow the use of one
step size to provide optimal accuracy for both derivative approximations.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{R. Abreu, D. Stich, and J. Morales (2013)}
\subsection{Abstract of prior work in field}

We generalize the well known Complex Step Method for computing derivatives by
introducing a complex step in a strict sense. Exploring different combinations
of terms, we derive 52 approximations for computing the first order derivatives
and 43 for the second order derivatives. For an appropriate combination of
terms and appropriate choice of the step size in the real and imaginary
directions, fourth order accuracy can be achieved in a very simple and
efficient scheme on a compact stencil. New different ways of computing second
order derivatives in one single step are shown. Many of the first order
derivative approximations avoid the problem of subtractive cancellation
inherent to the classic finite difference approximations for real valued steps,
and the superior accuracy and stability of the generalized complex step
approximations are demonstrated for an analytic test function.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{J. N. Lyness and C. B. Moler (1967)}
\subsection{Abstract of prior work in field}

Many algebraic computer languages now include facili- ties for the evaluation
of functions of a complex variable. Such facilities can be effectively used for
numerical differentiation. The method we de- scribe is appropriate for
computing the derivatives f(n) (x) of any analytic function which can be
evaluated at points in the complex plane near x.  Such a function might be a
complicated rational combination of elementary functions of x which is
difficult to differentiate analytically.  In this paper we derive several
formulas, any of which is suitable for evaluating the nth derivative of a
complex analytic function at a point in terms of function evaluations at
neighborimg points. For simplicity we assume throughout this paper that the
derivatives are to be evaluated at the origin. In this Introduction we state
one formula; in succeeding sections we prove this and related formulas, discuss
their degree and the error in their application, and give an example.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{J. N. Lyness (1968)}
\subsection{Abstract of prior work in field}

In a previous paper (Lyness and Moler [1]), several closely related formulas of
use for obtaining a derivative of an analytic function numerically are derived.
Each of these formulas consists of a convergent series, each term being a sum
of function evaluations in the complex plane.  In this paper we introduce a
simple generalization of the previous methods; we investigate the "truncation
error" associated with truncating the infinite series.  Finally we recommend a
particular differentiation rule, not given in the previous paper.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\end{document}