### 2014-02-27 response/response.tex
more progress on II

\documentclass{article}
\usepackage[left=1.0in,right=1.0in,top=1.0in,bottom=1.0in]{geometry}
\usepackage{fancyhdr}
\usepackage{color}
\usepackage{amsmath}
\definecolor{light-gray}{gray}{0.5}

\begin{document}

\begin{center}
    {\LARGE \bf Point-by-Point Response to Reviewer's Comments} \\
     Author's Responses in {\color{red} Red}
\end{center}

%%%% First Reviewer %%%%
\section{Re: Information to the authors}
%
\subsection*{Larger issues}
%
\begin{enumerate}
    \item
         Some parts of the manuscript suggest that the authors introduced a new method,
         i.e., the complex-step (CS) method. For example, in the introduction, they say
         “A goal of this study was to develop and evaluate a new, accurate, and practical
         method ...” Another example appears in page 3, where the authors say “The aim
         of this paper is not only to introduce the new complex-step method in the context
         of evaluating tangent-stiffness matrices ..”. However, the CS method has already
         appeared in many publications since 1998 [1]. Please revise the way you refer to
         the CS method along the manuscript

{\color{red}
  \begin{itemize}
      \item
        The way the CS method is referred to has been revised throughout,
        specifically: 
    \item In the abstract, "... the new method ..." has been
        replaced by "... the method ...", "... newly implemented ..." has been
        replaced by "... implemented ..."; 
    \item in the introduction, paragraph two,
        "... a new, accurate ..." has been replaced by "... a relatively
        underexplored, accurate ...", in paragraph two and four "... the new
        method ..." has been replaced by "... the method ...", in paragraph
        five, "... introduce the new complex-step method in the context of ..."
        has been replaced by "... study the complex-step method of ...", "...
        the new complex-step ..." has been replaced by "... the complex-step
        ...".
    \item 
        It is the authors' opinion that the above changes correct the suggestion
        that the authors introduce a new method.

  \end{itemize}}

    \item
        The work presented focuses on a parallel computational peridynamic code, and
        uses peridynamic models for mechanics. However, no detailed explanation of the
        equations involved is presented. Although the tangent-stiffness calculations are
        not specialized to Peridigm, the authors should include a short section introducing
        the peridynamic theory and the specific constitutive model used, i.e., the elastic
        peridynamic solid. You may add this to the already included very brief description
        of peridynamics in page 9 (lines 15-26). This will also help to explain the concept
        of the peridynamic horizon in page 10.

{\color{red}
  \begin{itemize}
    \item 
        We have revised the brief description of peridynamics in page 9 to be more complete:
    \item
        Material from silling2007 is interpreted including: the definition of vector states,
        the derivation of conservation laws in terms of state-based peridynamics, the definition
        of the isotropic elastic solid AKA the Linear elastic Peridynamic Solid.
    \item
        It is the authors' opinion that the additional material mentioned above explains the equations involved
        in the peridynamic models for mechanics, used in relation to the peridynamics code
        studied in the work.

  \end{itemize}}
    \item 
        The manuscript seems to be written in a language more familiar to computer scientists. 
        I would encourage the authors to revise certain parts of the manuscript to be
        presented in a language more familiar to the computational mechanics community.

{\color{red}
  \begin{itemize}
    \item 
        We have revised certain parts of the manuscript for clarity, adding additional information for the reader's
        convenience:
    \item
        In subsection 'Implementing the complex-step method in Peridigm', explanation of what is meant by datatype,
        container class, template, scope and instantiation are included since these concepts may be unfamiliar to
        a computational mechanics audience at least as denoted by these terms. 
    \item 
        In subsection 'The Comparative Study', the way message passing computations can slow runtime is explained
        by an analogy to an experiment given as an exercise for the reader. 
    \item 
        In section 'Conclusions', the term byte-copiable is explained. 
    \item 
        It is the authors' opinion that the above revisions to the language of the manuscript allow the intended audience
        to follow the discussion where concepts from computer science are relevant.
  \end{itemize}}

  \item
     In Section 1.1.1, please include a more comprehensive literature review for
     the CS method, including further references (some of them are listed below).
     In particular, specify what aspects of the CS method have been investigated in
     each reference. Then, you need to clearly mention your contribution within
     the CS method, in contrast to the other references (some comments in this
     regard appear in the manuscript in page 8).

{\color{red}
  \begin{itemize}
     \item 
         The authors have appended a more comprehensive literature review for the CS method to section 1.1.1.
         The aspects of CS investigated by the prior work, as well as any particularly interesting
         developments introduced by those works, have been listed. To argue for the novelty of the instant work,
         each prior work is compared to the instant work in terms of the use of the CS method and the similarity of the fields of research of the works.
     \item
         The literature review concludes with an argument for the novelty of the instant
         work where the prior works are taken as a whole and compared to the instant work. It is the authors' opinion
         that the instant work is novel for at least the reasons set out in the literature review material, where
         the contribution to the CS method is in the comparison to other methods in terms of computational
         time,  parallel implementation, and implementation within \emph{Peridigm} .
  \end{itemize}}

  \item
      The comparisons of the different algorithms are only performed with respect to the
      construction of tangent-stiffness matrices. In particular, the authors always use the
      solution of AD at each time step to compute the entries of the tangent-stiffness
      matrices, even for the FD, CD, and CS methods. However, the authors need to
      demonstrate the performance of the CS method, in comparison to the other ones,
      with respect to the solution of real, ideally non-linear problems.

{\color{red}
  \begin{itemize}
    \item 
        To address the reviewers comment, material from work done after the
        initial submission of the manuscript has been included. New sections
        include 'Working examples of CS, AD and FD for Newton's method' with
        new subsections. To existing sections were appended new subsections
        including 'Convergence and displacement accuracy studies', 'For further
        studies of Jacobian accuracy in Newton's method', 'Grain of salt for
        speed results' and 'Remarks'. Specifically, the added material addresses
        the reviewers comments:
    \end{itemize}}

\item 
      In other words, how the solutions compare in terms of accuracy and speed between the methods when
      solving given problems? Please include two examples of interest. The authors
      suggest that two examples involving non-linear systems are already included in the
      paper repository.

{\color{red}
  \begin{itemize}
    \item 
        The added section 'Working examples of CS, AD and FD for Newton's method' includes
        the two examples alluded to in the manuscript and requested by the reviewer. 
        % 
        To summarize, in this section is an example of interest: a displacement accuracy study and 
        a convergence rate study on a non-linear beam problem, where each of the methods
        discussed in the manuscript are used individually to solve the problem. 
 \end{itemize}}

\item
      The authors say in the Conclusion: “While the results showed that CS
      produced more accurate tangent-stiffness matrices than CD and FD under
      the parameters of the tests, it was not determined whether or not this is
      a clear advantage of CS over FD in terms of accuracy of final predicted
      displacements and speed of convergence.” This needs to be studied and
      results in this regard added to this manuscript.

{\color{red}
  \begin{itemize}
    \item 
      The added subsection 'Convergence and displacement accuracy studies'
      includes convergence rate and displacement accuracy studies similar to
      those shown in 'Working examples ...', but with the models and tools of
      the main study. This subsection provides an example of interest. 
  \item 
      To summarize, these studies are done in \emph{Peridigm} using the most refined mesh
      from the 'serial' test runs from the original manuscript. Measured was the performance of the methods in terms of speed
      (convergence rate) and accuracy (displacement accuracy).  Iteration speed
      was reported in the original manuscript.
  \item
      It was determined that the accuracy of CS over FD resulted in no appreciable
      advantage in terms of displacement accuracy or convergence rate for the problems
      solved in this study. 
    \item
      Additional comment on the significance of the accuracy of the Jacobian
      to Newton's method is included in the subsections 'For further studies of
      Jacobian accuracy in Newton's method', 'Grain of salt for speed results'
      and 'Remarks'. 
  \item 
      To summarize, Newton's method is insensitive to the accuracy differences
      shown by the methods, for the problems shown, when the methods are not
      deliberately misused with inappropriate step sizes. For completeness,
      formulas describing a relationship between Jacobian accuracy and
      convergence rate order are included. 
  \item
      It is the authors' opinion that the above revisions offer the discussion of relative problem solving
      performance of the methods that the reviewer has requested, including consideration
      of speed and accuracy in solving real, non-linear problems.
  \end{itemize}}
  \end{enumerate}

  \subsection*{Smaller issues}
  \begin{enumerate}

  \item
	Pg. 2: In line 14, explain what does the term “algorithmically consistent” mean.

{\color{red}
  \begin{itemize}
     \item
      A brief definition for the convenience of the reader was added in the manuscript. But, algorithmically consistent is a fairly common term in plasticity/mechanics. This same meaning is used for 'consistent' refering to 'tangent matrices' in \cite[perez2012consistent, perez2010]. It means consistent with integrating the plasticity model under the Khun-Tucker constraints. Integrating the plasticity model means, given a strain increment, determining a stress that does not exceed the yield stress, and by doing so computing a linear slope, called a consistent tangent modulus, that relates the strain increment to a correct stress increment. An alogorithmically consistent tangent stiffness matrix is then one that uses the consistent tangent modulii, making it a more accurate approximation of the stiffness of the system. 
  \end{itemize}}

  \item
	Pg. 2: In line 36, what do you mean by “discrimination”?

{\color{red}  
\begin{itemize}
     \item
      The word, 'discrimination' has been removed from the manuscript. 'Discrimination' was meant to mean 'measuring differences and making judgements on relative usefulness'. Since this meaning is probably understood from the word 'comparison' alone, the word 'discrimination' has been removed from the manuscript for clarity.
  \end{itemize}}

` \item	
	Pg. 2: In line 37, explain what do you mean by “in-situ instrumentation”.

{\color{red}  
\begin{itemize}
     \item
      The words, 'in-situ' have been removed from the manuscript. 'In-situ' was meant to suggest that timing measurements were taken from within and by the program being analyzed itself, rather than by an external program, but it has been removed for clarity.
  \end{itemize}}

  \item 
	Pg. 3: In lines 19-29, specify the sections where each part of the manuscript is
included.

{\color{red}  
\begin{itemize}
     \item
      The correction is made in the manuscript.
  \end{itemize}}

  \item 
Pg. 3: You may consider including the exact web address of the paper repository.

{\color{red}  
\begin{itemize}
     \item
      The correction is made in the manuscript. An exact web address for the paper repository has been added.
  \end{itemize}}

  \item 
Pg. 3: Please include the forward and centered ﬁnite difference expressions for
easier comparison with (2); include the truncation error as well. You do not have to
derive these expressions though.

{\color{red}  
\begin{itemize}
     \item
      The correction is made in the manuscript. Expressions for forward and centered difference have been included for faster comparison with (2).  
  \end{itemize}}

  \item 
Pg. 4: In Eq. (1), “$\partial f^2$” should be “$\frac{\partial^2 f}{\partial x^2}$”; similarly, for the third derivative.

{\color{red}  
\begin{itemize}
     \item
      The correction is made in the manuscript.
  \end{itemize}}

  \item 
Pg. 4: You may consider writing Eq. (2) with the truncation error

{\color{red}  
\begin{itemize}
     \item
      The correction is made in the manuscript.
  \end{itemize}}

9. Pg. 6: In lines 9-11, please add references to the “Poisson problem” and to the
“mathematical optimization”.

10. Pg. 7: Should Eqs. (4), (5), and (6) use “$\approx$” instead of “$=$”?

11. Pg. 7: In line 28, note that (cid:126)u is not the current conﬁguration, but the displacement
ﬁeld. Please revise the sentence.

12. Pg. 7: Please provide an expression for ${F^int}_i$
in the model used.

13. Pg. 9: In line 26, please add the reference [12] below to “coupling to molecular
simulations [29].”

14. Pg. 10: Please add an illustration of the block undergoing tension in Section 2.2,
for clarity. Also, clarify the displacement boundary conditions; are they applied in a
volumetric layer?

15. Pg. 10: In line 40, you mentioned a bandwidth of 7. This seems to be clear in 1D.
Does this hold in higher dimensions? Please clarify.

16. Pg. 10: In lines 47-49, you say “The speciﬁc parameters ... can be found ... in the
paper repository.” It seems a good idea to have the paper repository with ﬁles that a
user can eventually use. However, the content of the paper should be independent
of the repository. Please include a table with the parameters used. Similarly, in
page 11 (lines 11-13).

17. Pg. 14: In line 55, is there any conclusion you can draw from the power law index
$1.0 * 10^-6$?
{\color{red}  
\begin{itemize}
     \item
      Manuscript updated. More significant than the power law index is that the methods have the same power law relationship with computation time at all, meaning that there is some, but weak evidence that the methods have equal algorithm complexity. 
  \end{itemize}}

18. Pgs. 14-: Please clarify the value of the meshsize used on each of the experiments.
{\color{red}  
\begin{itemize}
     \item
      [TODO]
  \end{itemize}}

19. Pg. 16: Would it be possible to rescale Fig. 2, so that the dependence of the time
on the number of cores will look indeed linear?
{\color{red}  
\begin{itemize}
     \item
     Manuscript is in error and has been corrected. An accurate statement is that CPU time and number of cores are inversely related.
  \end{itemize}}

20. Pg. 16: In lines 54-56, you say “It should be pointed out that compared to the (cid:96)2 -
norms of any of the TS matrices by themselves, the magnitude of the accuracy
metric ... is relatively small.” Please include in the manuscript the explicit informa-
tion on the relative error. It seems that you already have such information in the
paper repository.
{\color{red}  
\begin{itemize}
     \item
[TODO]
  \end{itemize}}

21. Pg. 17: Include the values of the probe distances h used for each method.
{\color{red}  
\begin{itemize}
     \item
[TODO]
     The probe distances used in Peridigm are automatically determined with a heuristic algorithm and are not set by the user. They are 1.E-6 times the node spacing [ADAGIO REF] for FD and CD which is determined to limit the onset of subtractive cancellation. 1.E-100 was hard-coded for CS, while AD had no probe distance.
  \end{itemize}}

22. Pg. 17: In lines 53-54, you say “Accuracy trends matched those for the single core
tests.” Indeed, comparing Figs. 3 and 4, the results for FD seem to be quite the
same in serial and in parallel. For CS the accuracy seems higher in serial. For CD,
the accuracy is also higher in serial and, in serial, it has a ﬂuctuating behavior. Can
you comment on these observations?
{\color{red}  
\begin{itemize}
     \item
[TODO]
	The fluctuation seems to be a function of the amount of nonzero TS matrix elements relative to total TS matrix elements. Information supporting this has been added to the manuscript. This may be related to the fact that neighborhood radius in terms of number of Peridynamic nodes was kept constant and may indicate that connectivity in the mesh oscillates when the mesh is refined incrementally. 
  \end{itemize}}

23. Pgs. 17-18: What are the units of K ? If F int in Eq. (4) etc. is internal force, then K
should have units of force/length. However, in Figs. 3 and 4 you use the unit of
MPa, which is force/length2 . Please clarify.

24. Pg. 19: Are the units of the vertical axis in Fig. 5 “seconds”, as in Figs. 1 and 2?
Please indicate that. Also, is the label of the vertical axis in Fig. 6 correct?

25. Pg. 20: In lines 32-35, clarify what “byte-copyable” mean and why is that important.
{\color{red}  
\begin{itemize}
     \item
     Clarification made in manuscript, new reference added. 'Byte-copiable' means that members of an array of data elements are all of the same size in memory so that if one knew the position of the first element in the array, the position of all other elements in the array could be inferred with the additional information of the length of the array. This is important, in particular, to the designers and users of the thousands of MIC accelerator cards installed in the \emph{Stampede} computer system. MIC accelerator cards permit only the transfer of byte-copiable data from the host computer system for use with the '#pragma offload' mode, which is to say most code written for the MIC. This is even more important when it is realized that the great majority of \emph{Stampede}'s theoretical processing power is counted from these MIC accelerator cards. Meaning that if an AD user wanted to use any of that power, they would be out of luck since not even the full time personel at TACC could manage to compile the AD library Trilinos::Sacado for the MIC. A more 'casual' computational scientist would find the task impossible.
  \end{itemize}}

26. Pg. 21: Please include a section title for the references.

27. Pg. 24: In line 23, you may add the truncation error O(h2 ) to the expression. Oth-
erwise, you may need to replace the “=” sign by the “≈” sign. However, adding
the error term would make more clear the limiting expression when h → 0. Also,
replace the comma by a period.

28. Pg. 24: Why do we need S and X in Appendix B?
{\color{red}  
\begin{itemize}
     \item
     All the members of set X need to be members of set S otherwise the approximations don't work. X are numbers from the 'real' world and S are numbers from the computer world. For example: "http://en.wikipedia.org/wiki/IEEE_floating_point" shows that computers and languages following IEEE standards don't represent very large numbers with very fine precision at the same time due to memory constraints, and there are upper and lower bounds on the magnitudes of nonzero numbers that can be represented. Obviously, approximation error is expected, but the fact that a user doesn't actually get 'all' the numbers is something that is not obvious.
  \end{itemize}}

29. Pgs. 25-26: Are the functions in the second column of the table in 2(c) (page 25)
approximations? Why do you label the column “Approximates Function”? Simi-
larly, in line 16 (page 26) you say “approximated by the partial derivative”; is it an
approximation?
{\color{red}  
\begin{itemize}
     \item
      A computer can only provide finite precision approximations, of approxmations, of functions. Approximations of derivatives of functions are also approximations, therefor the label is correct in both cases. For example: "https://sourceware.org/git/?p=glibc.git;a=blob;f=sysdeps/x86_64/fpu/e_expl.S" is the implementation of the $exp(x)$ function on 80-bit extended precision floating point numbers in the popular GCC compiler. This implementation is the most accurate that GCC provides for the $e$ function on real numbers, but ultimately is still a discrete numerical algorithm.
  \end{itemize}}

30. Pg. 26: Please clarify whether storage is needed in automatic differentiation. For
example, do we need to store w, v , and u?

{\color{red}  
\begin{itemize}
     \item
      Clarification is added in the manuscript. For an actual AD system like Trilinos::Sacado or ADIFOR, w, v and u (stand-ins for simple or elementary math operations) need to be written in code differently from the built-in implementation to replace all math operations as part of a complete AD datatype. What doesn't absolutely have to be stored are the evaluations of w, v and u. Not storing the evaluations happens in 'forward' AD, storing the evaluations happens in 'backward' or 'reverse accumulation' AD.  
  \end{itemize}}

  \end{enumerate}
\end{document}